Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-26 paper/paper.tex
Edit provisioning delays section

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
    \caption{Percentage of high scores achieved by each algorithm for pareto 
    unsorted CyberShake ensembles when task runtime is scaled by a constant 
    factor. A scaling factor of $x$ means that the runtime of tasks in each 
    workflow was multiplied by $x$.}
    \label{fig:stretching}
\end{figure}

@@

\begin{figure*}[tb]
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
    }
    \caption{Boxplots for budget and deadline ratios when runtime estimate 
    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. A 
    value of 1 indicates that the actual cost is equal to the budget, or 
    the actual makespan is equal to the deadline, respectively. Values 
    greater than 1 indicate that the constraint has been exceeded.}
    \label{fig:variances}
\end{figure*}

@@

One important issue to consider when provisioning resources in the cloud is the amount of time between when a resource is requested, and when it becomes available to the application. Typically these provisioning delays are on the order of a few minutes, and are highly dependent upon the architecture of the cloud system and/or the size of the user's VM image~\cite{Nurmi2008b}.

Recall that it is assumed that resources are billed from the minute that they are requested until they are terminated, and not from the moment that they become available. As a result, provisioning delays have an impact on both the cost (budget constraint) and makespan (deadline constraint) of the ensemble.

Figure~\ref{fig:delays} shows the ratios of actual values to constraints when the provisioning delay is increased from 0 seconds up to 15 minutes. The simulations were done using uniform unsorted ensembles of 50 workflows from the Montage application. Each plot in the figure contains three boxes: one for each algorithm. Each box summarizes the results of 10 budgets, 10 deadlines, and 10 random seeds for a total of 1000 simulations per algorithm and provisioning delay (3000 simulations per plot). As in Figure~\ref{fig:variances}, the y-axis in each plot represents the ratio of the actual simulated value to the constraint value for the budget constraint (actual cost) or the deadline constraint (actual makespan). These ratios indicate how frequently and by how much the actual simulated values exceeded (or did not exceed) the constraints.

%\begin{figure}[ht]
%  \centering
%  \includegraphics[width=0.5\textwidth]{variances_scores}
%  \caption{Percentage of high scores achieved by each algorithm on Montage with
%  uniform unsorted ensembles when runtime estimate error varies from $\pm$0\%
%  to $\pm$50\%.}
%  \label{fig:variances_scores}
%\end{figure}

\begin{figure*}[tb]
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{    
        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
    }
    \caption[Boxplots for budget and deadline ratios when provisioning delays occur]{Boxplots for budget and deadline ratios when provisioning delay varies from 0 seconds to 15 minutes for all three algorithms. A value of 1 indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively. Values greater than 1 indicate that the budget/deadline has been exceeded.}
    \label{fig:delays}
\end{figure*}

The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not. The explanation for this behavior is also similar. The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget. Figure~\ref{fig:delays}.a shows that even for an impossibly short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget. In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget. With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.

% Figure~\ref{fig:delays_scores} shows the relative performance as a percentage of the number of high scores achieved by each algorithm as the provisioning delay increases. This figure shows that, as the delay increases, the relative performance of the static algorithm increases as well. This behavior is a result of the fact that the static algorithm is, in essence, cheating by using more time and money than the dynamic algorithms. In the previous section we showed how the relative performance of the algorithms remains the same when a uniform error is applied to the runtime constraints. In that case, the dynamic algorithms adapted to the error by rearranging the schedule to accomplish the same amount of work given the same deadline and budget. In this case, the dynamic algorithms adapt by performing less work (executing fewer workflows) to remain within the constraints while accounting for the delays. In both cases, the static algorithm completed the same amount of work, but did so by exceeding the constraints.

% \begin{figure}[ht]
%   \centering
%   \includegraphics[width=0.5\textwidth]{delays_scores}
%   \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when provisioning delay varies from 0 seconds to 15 minutes.}
%   \label{fig:delays_scores}
% \end{figure}

@@

\begin{figure*}[tb]
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-cost-all}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{
        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-makespan-all}
    }
    \caption{Boxplots for budget and deadline ratios when failure rate varies
    from 0 seconds to 50 \% for all three algorithms. A value of 1
    indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively. Values greater than 1 indicate that the budget/deadline has been exceeded.}
    \label{fig:failures}
\end{figure*}

@@

% This balances the columns by inserting a column break at the given reference
\IEEEtriggeratref{23}

>>>>

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
    \caption{Percentage of high scores achieved by each algorithm for pareto 
    unsorted CyberShake ensembles when task runtime is scaled by a constant 
    factor. A scaling factor of $x$ means that the runtime of tasks in each 
    workflow was multiplied by $x$.}
    \label{fig:stretching}
\end{figure}

@@

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
    }
    \caption{Boxplots for budget and deadline ratios when runtime estimate 
    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:variances}
\end{figure*}

@@

One important issue to consider when provisioning resources in the cloud is
the amount of time between when a resource is requested, and when it actually
becomes available to the application. Typically these provisioning delays are
on the order of a few minutes, and are highly dependent upon the architecture
of the cloud system and/or the size of the user's VM image~\cite{Nurmi2008b}.

We assume that resources are billed from the minute that they are requested
until they are terminated, and not from the moment that they become available.
As a result, provisioning delays have an impact on both the cost (budget
constraint) and makespan (deadline constraint) of the ensemble.

\TODO{Update the following paragraph depending on what results Maciek included}

Figure~\ref{fig:delays} shows the ratios of actual values to constraints when
the provisioning delay is increased from 0 seconds up to 15 minutes. These
simulations were done using 10 uniform unsorted ensembles of 50 workflows from
the Montage application, using 10 deadlines (1000 simulations per algorithm
and provisioning delay). As in Figure~\ref{fig:variances}, the y-axis in each
plot represents the ratio of the actual simulated value to the constraint
value for the budget constraint (actual cost) or the deadline constraint
(actual makespan). These ratios indicate how frequently and by how much the
actual simulated values exceeded (or did not exceed) the constraints.

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{    
        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
    }
    \caption{Boxplots for budget and deadline ratios when provisioning delay 
    varies from 0 seconds to 15 minutes for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:delays}
\end{figure*}

\TODO{Update this next paragraph depending on what Maciek included in the figure. The results seem to have changed.}

The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not. The explanation for this behavior is also similar. The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget. Figure~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget. In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget. With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.

% Figure~\ref{fig:delays_scores} shows the relative performance as a
% percentage of the number of high scores achieved by each algorithm as the
% provisioning delay increases. This figure shows that, as the delay
% increases, the relative performance of the static algorithm increases as
% well. This behavior is a result of the fact that the static algorithm is, in
% essence, cheating by using more time and money than the dynamic algorithms.
% In the previous section we showed how the relative performance of the
% algorithms remains the same when a uniform error is applied to the runtime
% constraints. In that case, the dynamic algorithms adapted to the error by
% rearranging the schedule to accomplish the same amount of work given the
% same deadline and budget. In this case, the dynamic algorithms adapt by
% performing less work (executing fewer workflows) to remain within the
% constraints while accounting for the delays. In both cases, the static
% algorithm completed the same amount of work, but did so by exceeding the
% constraints.

% \begin{figure}[ht]
%   \centering
%   \includegraphics[width=0.5\textwidth]{delays_scores}
%   \caption{Percentage of high scores achieved by each algorithm on Montage 
%   with uniform unsorted ensembles when provisioning delay varies from 0 
%   seconds to 15 minutes.}
%   \label{fig:delays_scores}
% \end{figure}

\TODO{Update this paragraph when Maciek explains what was included in the plot.}

@@

\begin{figure*}[tb]
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-cost-all}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{
        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-makespan-all}
    }
    \caption{Boxplots for budget and deadline ratios when failure rate varies
    from 0\% to 50\% for all three algorithms. Values greater than 1 indicate 
    that the budget/deadline constraint was exceeded.}
    \label{fig:failures}
\end{figure*}

@@

% This balances the columns by inserting a column break at the given reference
\IEEEtriggeratref{19}
